---
title: "Homework 6"
author: "Yunshen Bai"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```



```{r load_libraries,message=FALSE,warning=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(MASS)
library(dplyr)
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("./data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  dplyr::select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(formula=resolution ~ victim_age + victim_sex + victim_race, family = binomial())

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  dplyr::select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  dplyr::select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  dplyr::select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

For this problem, we’ll use the Central Park weather data. The code chunk below will download these data.
```{r,message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  dplyr::select(name, id, everything())
```

The function to generate our bootstrap samples is
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of $R^2$ and $log(\beta_0*\beta_1)$.

```{r}
boot_straps_weather = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df|>drop_na()))
  )

```

Next, we use `broom::glance()` and `broom::tidy()` functions to get these two quantities.

```{r}
bootstrap_results_log_beta0_beta1 = 
  boot_straps_weather |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin+prcp, data = df) ),
    results = map(models, broom::tidy)) |> 
  dplyr::select(-strap_sample, -models) |>
  unnest(results) |>
  dplyr::select(strap_number,term,estimate)|>
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  )|>
  janitor::clean_names()|>
  mutate(log_beta1_beta2=log(tmin*intercept))|>
  pull(log_beta1_beta2)
```

```{r}
bootstrap_results_r2 = 
  boot_straps_weather |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin+prcp, data = df) ),
    results = map(models, broom::glance)) |> 
  dplyr::select(-strap_sample, -models) |>
  unnest(results) |>
  pull(r.squared)
```

Then, we plot the distribution of these two quantities, and describe these in words

```{r}
ggplot(data.frame(bootstrap_results_r2),aes(x=bootstrap_results_r2))+
  theme_bw() + 
  geom_density()+
  labs(
    x="R^2",
    title = "Distribution of R^2"
  )
```


```{r}
ggplot(data.frame(bootstrap_results_log_beta0_beta1),aes(x=bootstrap_results_log_beta0_beta1))+
  theme_bw() + 
  geom_density()+
  labs(
    x="log(beta0*beta1)",
    title = "Distribution of log(beta0*beta1)"
  )
```

From above two density plots the distribution of $R^2$ and $log(\beta_0*\beta_1)$, they under the normal distribution. Then, we are interested in the confidence interval of these two quantities. We use the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for $R^2$ and $log(\beta_0*\beta_1)$.

```{r}
CI_R2=quantile(bootstrap_results_r2,c(0.025,0.975))
CI_log_beta0_beta1=quantile(bootstrap_results_log_beta0_beta1,c(0.025,0.975))
CI_R2
CI_log_beta0_beta1
```
Therefore, the 95% confidence interval for $R^2$ is (`r CI_R2`), the 95% confidence interval for $log(\beta_0*\beta_1)$ is (`r CI_log_beta0_beta1`).

## Problem 3

In this problem, we will analyze data gathered to understand the effects of several variables on a child’s birthweight. Then, load and clean the data for regression analysis

```{r,message=FALSE}
birthweight=
  read_csv("./data/birthweight.csv")|>
  drop_na()|>
  mutate(
    babysex=factor(babysex),
    frace=factor(frace),
    malform=factor(malform),
    mrace=factor(mrace)
  )
```
This dataset consists of `r nrow(birthweight)` children and `r ncol(birthweight)` variables, and four of variables are categorical.

Then, propose a regression model for birthweight. First step, we try the full model and see its performance.

```{r}
full_model=lm(bwt~.,data = birthweight)
summary(full_model)
```

Then, we perform stepwise model selection by AIC to select variables.
```{r}
step_model=stepAIC(full_model)
```

```{r}
summary(step_model)
```


Form the result, we can know that some variable are not significant in this model. Therefore, we decide to drop these variables.

```{r}
step_model_update=update(step_model,.~.-fincome)
summary(step_model_update)
```
Next, we drop the outliers of this model.
```{r}
SR = stdres(step_model_update)
n = length(step_model_update$residuals) 
p = length(step_model_update$coefficients)
alpha = 0.05 
t.cutoff = qt(1- alpha/2, n-p)
outliers = which(abs(SR)>t.cutoff)
new_birthweight = birthweight[-outliers,]
drop_outliers_model=update(step_model_update,.~.,data=new_birthweight)
summary(drop_outliers_model)
```



Some measurement of this model are
```{r}
data.frame(R_square_adjusted=summary(drop_outliers_model)$adj.r.squared,
           AIC=AIC(drop_outliers_model),
           BIC=BIC(drop_outliers_model),
           PRESS=MPV::PRESS(drop_outliers_model))|>
  knitr::kable()
```

Next, we will show a plot of model residuals against fitted values 
```{r}
pred=
  new_birthweight|>
  add_predictions(drop_outliers_model)|>
  pull(pred)
res=drop_outliers_model$residuals
ggplot(aes(x = pred, y = res),data=data.frame(pred,res)) + 
  geom_point() + 
  geom_smooth(color="red")
```
Then, we compare our model to two others:

- One using length at birth and gestational age as predictors (main effects only)
- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
model2=lm(bwt~blength+gaweeks,data = birthweight)
model3=lm(bwt~bhead+blength+babysex+bhead:blength+bhead:babysex+blength:babysex+bhead:blength:babysex,data = birthweight)
```

Then, we perfome cross validation to compare these models.

```{r}
cv_df = 
  crossv_mc(birthweight, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))|>
  mutate(
    drop_outliers_model  = map(train, \(df) drop_outliers_model),
    model2  = map(train, \(df) model2),
    model3  = map(train, \(df) model3)) |> 
  mutate(
    rmse_drop_outliers_model = map2_dbl(drop_outliers_model, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model2 = map2_dbl(model2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_model3 = map2_dbl(model3, test, \(mod, df) rmse(model = mod, data = df)))
```

Then, we drow a plot to compare the rmse of these three model

```{r}
cv_df |> 
  dplyr::select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From this plot, we can see that our final model has less rmse, which means that our model has a better performance on prediction.
